Daft-Introduction

### 2024-10-31 draftintro.Rmd
theorydraftfirst

---
title: "RQdraft2"
author: "Verooo"
date: "2024-09-18"
output: html_document
---

@@

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

@@

Collaboration has become an essential part of modern academia, with scholars increasingly engaging in joint research efforts. Co-authorship allows researchers to combine expertise, share resources, and enhance their scholarly output (Cheong & Corbitt, 2009). However, the dynamics of academic collaboration are not free from social patterns, and gender homophily—where individuals prefer to collaborate with others of the same gender—may play a significant role. Understanding how gender affects co-authorship networks, particularly in terms of homophily, can reveal important insights into how academic opportunities and achievements are distributed across genders.

Previous studies have found that male researchers often prefer collaborating with other men, while female researchers tend to collaborate with other women, albeit to a lesser extent (Horta et al., 2022). This raises critical questions about the structure and inclusiveness of academic networks. In fields such as Sociology and Political Science, where collaboration is crucial for scholarly success, gender-based patterns could shape the career trajectories of researchers. Furthermore, as Kwiek and Roszka (2021) point out, all-male authorship is far more prevalent than all-female authorship, indicating that male-dominated collaboration networks may reinforce existing gender imbalances in academia.

In this study, I explore the extent to which gender homophily influences co-authorship networks among faculty in the Sociology and Political Science Departments at Radboud University. Specifically, I aim to examine whether there are differences in gender homophily between the two departments, and how homophily patterns vary depending on the prestige of the journals in which researchers publish. Additionally, I investigate whether structural network factors—such as the size and composition of collaboration networks—affect gender homophily in each department.

The research will contribute to a growing body of literature that seeks to understand the social dynamics underlying academic collaboration. By analyzing gender homophily and network structures in two different academic disciplines, this study will shed light on how gender shapes collaboration opportunities and how these dynamics may differ across scholarly fields. Moreover, exploring the role of journal prestige in shaping homophily will offer insights into whether gender-based preferences shift in response to career-enhancing factors like publishing in high-prestige venues.

------------------------------------------------------------------------

# Research Questions

RQ1a: To what extent does gender homophily occur across all co-publications of the staff of the Sociology Department and Political Science Department at Radboud University?

RQ1b: To what extent do the two departments differ in their gender homophily?

RQ2: How to does gender homophily affect collaboration differently for collaborations between staff who have published in low vs. high prestige journals?

RQ3: To what extent do structural network effects influence the gender homophily of each department?

## Horta et al. (2022)

-   male researchers preferred to co-publish with other male researchers,

-   and female researchers preferred to collaborate with other females.

-   researchers of approximately similar ages tended to collaborate more with each other.

-   geographical, cultural, & institutional proximity were strongest predictors of collaboration.

-   analysis shows that in the social sciences, ACQUIRED attributes (particularly preferences regarding research agendas) take precedence over ASCRIBED attributes (e.g. gender or age)

-   Horta et al. (2022) - RQs:

    -

        (1) What homophily preferences determine research collaborations among higher education researchers?

    -

        (2) What are the most powerful homophily indicators in research collaborations in higher education research?

## Kwiek & Roszka (2021)

-   Most male scientists collaborate solely with males.
-   Most female scientists, in contrast, do not collaborate with females at all.
-   All-female authorship marginal, all-male authorship pervasive, in all age groups.
-   Articles written in mixed-sex collaboration published in more prestigious journals.

------------------------------------------------------------------------

# References:

Horta, H., Feng, S., & Santos, J. M. (2022). Homophily in higher education research: a perspective based on co-authorships. Scientometrics, 127(1), 523–543. <https://doi.org/10.1007/s11192-021-04227-z>

Kwiek, M., & Roszka, W. (2021). Gender-based homophily in research: A large-scale study of man-woman collaboration. Journal of Informetrics, 15(3), 101171. <https://doi.org/10.1016/j.joi.2021.101171>

# Data types needed for research

+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Variable                          | Potential measurement                                   | Notes                                                                                                  |
+===================================+=========================================================+========================================================================================================+
| **Actor attributes**              |                                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Name                              | Name from university department website                 | using "read_html()" function                                                                           |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Age                               | length of career                                        | time since first publication                                                                           |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Gender                            | Google Scholar:\                                        | What about gender neutral names? If we go by this principle, then gender neutral names = non-binary ?? |
|                                   | Name ➙ Look up in gendered list whether Woman or Man    |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
|                                   | Linked-in:\                                             | Not everyone puts their pronouns on linkedin\                                                          |
|                                   | Pronouns                                                | And: not everyone who uses e.g. She/her pronouns is a Woman                                            |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| High prestige publication (prior) | If the author had a publication in a journal that ....\ |                                                                                                        |
|                                   | before the publication of the paper in question         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Affiliation                       | Name of university they work at (radboud university)    | Taken from university website about department                                                         |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| **Relational Attributes**         |                                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Initiation of the co-authorship   | First author of paper = initiator                       |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Paper                             | A paper people have co-authorship on                    | How prevent from counting double?\                                                                     |
|                                   |                                                         | Overall: systematic approach to webscrape?                                                             |
|                                   | Signifier: DOI?                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
|                                   |                                                         |                                                                                                        |
+-----------------------------------+---------------------------------------------------------+--------------------------------------------------------------------------------------------------------+

: Data types

>>>>

---
title: "Daft-Introduction"
author: "Verooo"
date: "2024-10-31"
output: html_document
bibliography: references.bib
---

@@

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

Collaboration has become an essential part of modern academia, with scholars increasingly engaging in joint research efforts. Co-authorship allows researchers to combine expertise, share resources, and enhance their scholarly output (@Cheong2009). However, the dynamics of academic collaboration are not free from social patterns, and gender homophily—where individuals prefer to collaborate with others of the same gender—may play a significant role. Understanding how gender affects co-authorship networks, particularly in terms of homophily, can reveal important insights into how academic opportunities and achievements are distributed across genders.

Previous studies have found that male researchers often prefer collaborating with other men, while female researchers tend to collaborate with other women, albeit to a lesser extent (@Horta2022). This raises critical questions about the structure and inclusiveness of academic networks. In fields such as Sociology and Political Science, where collaboration is crucial for scholarly success, gender-based patterns could shape the career trajectories of researchers (@Cheong2009). Furthermore, as Kwiek and Roszka (@Kwiek2021) point out, all-male authorship is far more prevalent than all-female authorship, indicating that male-dominated collaboration networks may reinforce existing gender imbalances in academia.

Another influential factor in collaboration between scholars seems to be the scholars' 'prestige', namely whether they have published in renowned journals. **SOURCE** found that .....PART ABOUT JOURNAL PRESTIGE

In this study, I explore the extent to which gender homophily influences co-authorship networks among faculty in the Sociology and Political Science Departments at Radboud University. Specifically, I aim to examine whether there are differences in gender homophily between the two departments, and how homophily patterns vary depending on the prestige of the journals in which researchers publish.

This research will contribute to a growing body of literature that seeks to understand the social dynamics underlying academic collaboration. By analyzing gender homophily and network structures in two different academic disciplines, this study will shed light on how gender shapes collaboration opportunities and how these dynamics may differ across scholarly fields. Moreover, exploring the role of journal prestige in shaping homophily will offer insights into whether gender-based preferences shift in response to career-enhancing factors like publishing in high-prestige journals.

# References
